One-dimensional Classical Gas in a Harmonic Trap with Repulsive Interaction

### 2016-12-11 Zhiyu.tex
Dec11 PRE format, tex

\documentclass[aps,pre,twocolumn
,groupedaddress]{revtex4-1}
%\documentclass[a4paper,onecolumn,12pt]{article}
\usepackage{amsmath}
 \usepackage{amsfonts}
 \usepackage{amssymb}
 \usepackage{graphicx}
 \usepackage{subfigure}
 \usepackage{enumerate}
 \usepackage{verbatim}% for \begin{comment} \end{comment}
 \usepackage{color}
 \usepackage{bm}
   \usepackage[colorlinks,
            linkcolor=red,
            anchorcolor=blue,
            citecolor=green
            ]{hyperref}
 \begin{document}
 \title{One-dimensional Classical Gas in a Harmonic Trap with Repulsive Interaction}
 \author{Zhiyu}
 \affiliation{Fudan University, Shanghai, China}

\begin{abstract}
\end{abstract}
 \maketitle
% \tableofcontents
% \newpage
\section{Introduction}
In cold atom experiment, when BEC happens, atoms condense into their groud state. If one suddenly modify the trap potential or interaction parameter, which is just a ``quench", then the low lying collective modes will be excited. One of the most consipicuous collective phenomenon is the breathing mode. Breathing mode, also called as monopole mode, is periodic expanding and contracting behaviour of the cloud of atoms. The frequency of breathing mode is widely discussed in literature. When the atoms have interaction, it is observed that the frequncy is dependent on the interaction. xxx used mean-field technique to calculate the relation between breathing frequency and interaction[ ]. While xxx numerically calculated the frequency[ ]... BEC is a quantum gas, so the existing literature on breathing modes all focus on quantum gas. But the quantum property is not necessary to the breathing mode phenomenon, though the question of breathing mode generate from BEC. For classical gas, there is also such breathing behaviour. So a natural question would be asked is, how do the breathing frequency of classical gas dependent on interaction or other parameters. Surprisingly, to the best of our knowledge this question is not answered yet in the literature and thus need study.

Moreover, in a nonquilibrium many-body system, a more intriguing question is about its thermalization or relaxation property. Thermalization is among the most fundamental process for a many-body interacting system. All systems do not thermalize. An example is the Fermi–Pasta–Ulam paradox [ ] where shows confinement in phase-space so that the system stay non-thermalized for a long time. A lot of research on thermalization property has been done in other classical system. H De Raedt, et al. studied the one dimensional gravitational system with Lyapunov exponents[ ]. xxx studied the ring of harmonic oscillators and magnetic moment by examine the canonical distribution of a subsystem[ ].

In this paper, we answer these two questions. In the first part, we give an estimation of breathing, along with its explanation, and compare that with the numerical simulation result. We will also show a rotating frame method of treating this problem, which is a convenient tool to analyze simple harmonic systems. In the second part, we study the thermalization property. We first examine the canonical disribution of energy to find out the threshold of thermalization, i.e. under what parameter the thermalization time scale is very large. Then we point out that there are two kind of thermalization (or relaxation) behaviour that could be measured in this system. In the end we measure the Lyapunov Exponent to quantify the time scales and confirm the former discussion.

\section{Preparation}
\subsection{Scaling behaviour}
We investigated the classical gas in harmonic trap with a repulsive interaction. We chose one of the simplest types of interaction:
\begin{equation}
F=\left\lbrace
\begin{aligned}
&F=F0 	&,x\textless\sigma\\
&F=0 	&,x\textgreater\sigma
\end{aligned}
\right.
\label{eq:preparation1}
\end{equation}
so the Hamiltonian is
\begin{equation}
H=\frac{1}{2}m\omega_0^2\sum_{i}x_i^2+\frac{1}{2}m\sum_i v_i^2+\sum_{\left|x_i-x_j\right|\textless\sigma}F_0\left(\sigma-\left|x_i-x_j\right|\right)
\label{eq:preparation2}
\end{equation}
It can be shown that eq.\ref{eq:preparation2} can be rewrite into the following form:
\begin{equation}
\tilde{H}=\frac{1}{2}\sum_{i}\tilde{x_i}^2+\frac{1}{2}\sum_i\left(\frac{d\tilde{x}}{d\tilde{t}}\right)^2+\sum_{\left|\tilde{x_i}-\tilde{x_j}\right|\textless 1}F_0\left(1-\left|\tilde{x_i}-\tilde{x_j}\right|\right)
\end{equation}
with the transformation:
\begin{equation}
\left\lbrace
\begin{aligned}
&\tilde{H}&=&\frac{H}{m\omega_0^2\sigma^2}\\
&\tilde{x_i}&=&\frac{x_i}{\sigma}\\
&\tilde{t}&=&\omega_0t\\
&\tilde{F_0}&=&\frac{F_0}{m\omega_0^2\sigma}\\
\end{aligned}
\right.
\end{equation}

@@

\section{Breathing Frequency}
Usually, to see the response of quantum gas to a quench, one begin simulation with the ground state and then suddenly vary some parameter by a little, for instance, $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$. In that case, one will observe the radius R oscillating at a certain frequency, which is usually called ``breathing mode".
However, it is quite different in the classical version. Firstly, the ground state may be meaningless. In quantum case, the reason why we care about ground state and low lying excitations is that BEC occur at low temperature. But in classical case, there is no BEC. So it is not necessary to focus on low lying excitations in classical case. Besides, since classical limit is $\bar{h}\rightarrow 0$, no matter how low the kT is, $\bar{h}/kT$ is always zero so that classical behavious of low lying excitations is quite different from the quantum one. Secondly, it is no longer necessary to vary parameters ``by a little" in the quench. In quantum case, the small quench $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$ means the spectrum is shifted slightly, thus one may expect to see some beat phenomenon({\color{red}{?}}). But in classical case, since $\bar{h}$ vanishes, no matter how small the quench is, one can never recover or proximate the quantum case. Thus it is not necessary to limit ourselves to small quench.

@@

\section{Thermalization}
\subsection{Dynamics Study}
It is natural to think that a many-body system with interaction will be thermalized in ``usual" case. To study the condition of thermalization is equivalent to find out the mechanism that prohibits the system from ergodic. We begin the study with two-particle motion. 
\subsubsection{two particles}

@@

\newpage
\begin{figure}[hbtp]
\centering
%\includegraphics[scale=0.5]{D:/matlab2016/MolecularDynamics/figure/latexpic/distribution_at_N=5_F0=10000_pre.eps}
%\caption{Energy distribution, N=5, F0=10000}
%\label{fig:thermalization8}

@@

\begin{figure}
\subfigure[]{
\begin{minipage}[b]{0.3\textwidth}
\includegraphics[scale=0.17]{D:/matlab2016/MolecularDynamics/figure/latexpic/freq_scanE_1.png} 
\end{minipage}
}
\subfigure[]{
\begin{minipage}[b]{0.3\textwidth}
\includegraphics[scale=0.17]{D:/matlab2016/MolecularDynamics/figure/latexpic/freq_scanE_2.png} 
\end{minipage}
}
\subfigure[]{
\begin{minipage}[b]{0.3\textwidth}
\includegraphics[scale=0.17]{D:/matlab2016/MolecularDynamics/figure/latexpic/freq_scanE_3.png} 
\end{minipage}
}
\centering
\caption{Frequency of relative motion with internal energy in three regime} Numbers in the legend is the value of $\frac{F_0}{m\omega_0^2\sigma}$
\label{fig:thermalization3}
\end{figure}
\end{comment}

\newpage
\renewcommand\refname{Reference}
%\bibliographystyle{plain}
\bibliography{E:/Reference/mybib}

>>>>

\documentclass[aps,pre,twocolumn,groupedaddress]{revtex4-1}
%\documentclass[a4paper,onecolumn,12pt]{article}
%\documentclass[reprint, amsmath,amssymb,aps,]{revtex4-1}

\bibliographystyle{apsrev4-1}
\usepackage{amsmath}
 \usepackage{amsfonts}
 \usepackage{amssymb}
 \usepackage{graphicx}
 \usepackage{subfigure}
 \usepackage{enumerate}
 \usepackage{verbatim}% for \begin{comment} \end{comment}
 \usepackage{color}
 \usepackage{xcolor}
 \usepackage{bm}
 \usepackage{hyperref}
\hypersetup{
    colorlinks = false,
    linkbordercolor = {red},
    citebordercolor = {green},
}
 \begin{document}
 \title{One-dimensional Classical Gas in a Harmonic Trap with Repulsive Interaction}
 \author{Zhiyu}
 \affiliation{Fudan University, Shanghai, China}

\begin{abstract}
\end{abstract}
 \maketitle
% \tableofcontents
% \newpage
\section{Introduction}
Controlled and tunable experimental realizations of confined quantum systems with ultracold atoms enable nonequilibrium studies of quantum many-body states in novel geometries. A common feature of many cold-atom experiments is the confinement of a many-body system in a harmonic trap. Trapping introduces many distinctive features which have no analog in uniform many-body states, including collective excitations such as breathing modes, dipole modes, and scissors modes. Such trap-related collective modes have been widely studied both experimentally and theoretically for continuum systems, especially in the mean-field regime, since the early days after quantum degeneracy was achieved with trapped atoms \cite{Dalfovo1997}\cite{Jin1996}.

In mean-field perspective, the physics of atom gas in Bose-Einstein Condensation can be described by the Gross-Pitaevskii equation\cite{Dalfovo1999}\cite{Stringari1996}, where the breathing frequency for 1D Bose gas is calculated to be $\Omega_{GP}=\sqrt{3}\omega_0$. Meanwhile, the experiment of Ref.\cite{Haller2009} has found a regime of interactions where the breathing-mode frequency approaches the Gross-Pitaevskii prediction. However, since mean-field approach which generally omit quantum fluctuation is successful in this problem, one may wonder how quantum effect plays the role in the breathing mode. Thus, a natural question would be asked is how breathing frequency is dependent on interaction for classical gas. Though generated in BEC, the breathing mode phenomenon is not restricted to BEC. For a classical gas, the breathing mode still can be excited with a quench. A good knowledge of classical version may help us understand the quantum question better. Surprisingly,  to the best of our knowledge, all the literature focus on the breathing mode in BEC, without a careful study of their classical counterpart, which actually seems easier. Since this classical version question has not been answered yet, more study is needed.

One the other hand, in a nonquilibrium many-body system, a more intriguing question is about its thermalization or relaxation property. Thermalization is among the most fundamental process for a many-body interacting system. All systems do not thermalize. An example is the Fermi-Pasta-Ulam paradox [ ] which shows confinement in phase-space so that the system stay non-thermalized for a long time. A lot of research on thermalization property has been done in other classical system. Ref.\cite{Tsuchiya2000}\cite{Yawn1997} studied the one dimensional gravitational system with Lyapunov exponents. Ref.\cite{Jin2013} studied the ring of harmonic oscillators and magnetic moment by examine the canonical distribution of a subsystem . In our paper, we will also probe the system's thermalization time scales from the perspective of canonical distribution as wel as Lyapunov exponent. Wiht these approaches, we will show that the thermalization time scale manifest a competition between total energy and interaction energy, so there is not any singular behaviour in the thermalization time scales except the trivial divergence at zero interaction.

This paper is organized as follows: In Section \ref{section:preparation}, we perform a scaling to reduce the parameter of the system. In addition, we briefly observe the ground state and reveal two dynamically distinct regime of the system, i.e. strong interaction regime and weak interaction regime. In Section \ref{section:breathing frequency}, we evaluate the breathing frequency according to its mechanism, and compare that with our numerical simulation result. Besides, we also introduce a rotating frame method of treating this problem, which is a convenient tool to analyze simple harmonic systems. In section \ref{section:Thermalization}, we focus on the thermalization property. We first examine the canonical disribution of energy to find out the threshold of thermalization, i.e. under what parameter the thermalization time scale is very large. Then we point out that there are two kind of thermalization (or relaxation) behaviour that could be measured in this system. In the end of this section, we measure the Lyapunov Exponent to quantify the time scales and confirm the former discussion.

\section{Preparation}\label{section:preparation}
\subsection{Scaling behaviour}
We investigated the classical gas in harmonic trap with a repulsive interaction. We chose one of the simplest types of interaction:
\begin{equation}
F=\left\lbrace
\begin{aligned}
&F=F_0 	&,x\textless\sigma\\
&F=0 	&,x\textgreater\sigma
\end{aligned}
\right.
\label{eq:preparation1}
\end{equation}
so the Hamiltonian is
\begin{equation}
H=\frac{1}{2}m\omega_0^2\sum_{i}x_i^2+\frac{1}{2}m\sum_i v_i^2+\sum_{\left|x_i-x_j\right|\textless\sigma}F_0\left(\sigma-\left|x_i-x_j\right|\right)
\label{eq:preparation2}
\end{equation}
It can be shown that eq.\ref{eq:preparation2} can be rewrite into the following form:
\begin{equation}
\tilde{H}=\frac{1}{2}\sum_{i}\tilde{x_i}^2+\frac{1}{2}\sum_i\left(\frac{d\tilde{x}}{d\tilde{t}}\right)^2+\sum_{\left|\tilde{x_i}-\tilde{x_j}\right|\textless 1}F_0\left(1-\left|\tilde{x_i}-\tilde{x_j}\right|\right)
\end{equation}
with the transformation:
\begin{equation}
\left\lbrace
\begin{aligned}
&\tilde{H}&=&\frac{H}{m\omega_0^2\sigma^2}\\
&\tilde{x_i}&=&\frac{x_i}{\sigma}\\
&\tilde{t}&=&\omega_0t\\
&\tilde{F_0}&=&\frac{F_0}{m\omega_0^2\sigma}\\
\end{aligned}
\right.
\end{equation}

@@

\section{Breathing Frequency}\label{section:breathing frequency}
Usually, to see the response of quantum gas to a quench, one begin simulation with the ground state and then suddenly vary some parameter by a little, for instance, $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$. In that case, one will observe the radius R oscillating at a certain frequency, which is usually called ``breathing mode".
However, it is quite different in the classical version. Firstly, the ground state may be meaningless. In quantum case, the reason why we care about ground state and low lying excitations is that BEC occur at low temperature. But in classical case, there is no BEC. So it is not necessary to focus on low lying excitations in classical case. Besides, since classical limit is $\bar{h}\rightarrow 0$, no matter how low the kT is, $\bar{h}/kT$ is always zero so that classical behavious of low lying excitations is quite different from the quantum one. Secondly, it is no longer necessary to vary parameters ``by a little" in the quench. In quantum case, the small quench $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$ means the spectrum is shifted slightly, thus one may expect to see some beat phenomenon({\color{red}{?}}). But in classical case, since $\bar{h}$ vanishes, no matter how small the quench is, one can never recover or proximate the quantum case. Thus it is not necessary to limit ourselves to small quench.

@@

\section{Thermalization}\label{section:Thermalization}
\subsection{Dynamics Study}
It is natural to think that a many-body system with interaction will be thermalized in ``usual" case. To study the condition of thermalization is equivalent to find out the mechanism that prohibits the system from ergodic. We begin the study with two-particle motion. 
\subsubsection{two particles}

@@

\begin{figure}[hbtp]
\centering
%\includegraphics[scale=0.5]{D:/matlab2016/MolecularDynamics/figure/latexpic/distribution_at_N=5_F0=10000_pre.eps}
%\caption{Energy distribution, N=5, F0=10000}
%\label{fig:thermalization8}

@@

\begin{figure}
\subfigure[]{
\begin{minipage}[b]{0.3\textwidth}
\includegraphics[scale=0.17]{D:/matlab2016/MolecularDynamics/figure/latexpic/freq_scanE_1.png} 
\end{minipage}
}
\subfigure[]{
\begin{minipage}[b]{0.3\textwidth}
\includegraphics[scale=0.17]{D:/matlab2016/MolecularDynamics/figure/latexpic/freq_scanE_2.png} 
\end{minipage}
}
\subfigure[]{
\begin{minipage}[b]{0.3\textwidth}
\includegraphics[scale=0.17]{D:/matlab2016/MolecularDynamics/figure/latexpic/freq_scanE_3.png} 
\end{minipage}
}
\centering
\caption{Frequency of relative motion with internal energy in three regime} Numbers in the legend is the value of $\frac{F_0}{m\omega_0^2\sigma}$
\label{fig:thermalization3}
\end{figure}
\end{comment}
\cite{Dalfovo1997}
\cite{Tschischik2013}
\cite{Tschischik2015}
\cite{Guery-Odelin1999}
\cite{Yawn1997}
\cite{El-Zant1998}
\cite{Jin1996}
\cite{Tsuchiya2000}
\bibliography{mybib}
